---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figs/",
  cache.path = "cache/"
)
```

# Simulate from beta regressions with Stan or betareg

This example shows how to simulate from a beta regression model. First, I do this using the rstanarm package for a fully Bayesian model. Second, I do this with a maximum likelihood model fit with the betareg package by modifying the function `arm::sim()`.

# Packages

Beta regression was just added into rstanarm, so you will need to install the package from GitHub. This will require a [C++ compiler](https://support.rstudio.com/hc/en-us/articles/200486498-Package-Development-Prerequisites). Sometime in the near future I imagine you'll just be able to install the version from CRAN.

```{r}
# install.packages("devtools")
# run this once to install rstanarm!!
# devtools::install_github("stan-dev/rstanarm", args = "--preclean", build_vignettes = FALSE)

library(rstanarm)
options(mc.cores = parallel::detectCores())
library(betareg)
```

# Simulate data

First simulate some data to fit our models to. This is modified from the help file for `rstanarm::stan_betareg`.

```{r}
set.seed(1)
N <- 150
x <- rnorm(N, 2, 1)
mu <- binomial(link = "logit")$linkinv(0.8 + 0.3*x)
phi <- exp(1.5)
y <- rbeta(N, mu * phi, (1 - mu) * phi)
hist(y, col = "dark grey", border = FALSE, xlim = c(0,1))
fake_dat <- data.frame(y, x)
```

# Stan

Now we can fit the model with `rstanarm::stan_betareg`:

```{r}
fit <- stan_betareg(y ~ x, data = fake_dat, link = "logit")
print(fit)
fit$stanfit

plot(fit)
# pp_check(fit)
# prior_summary(fit)
```

We can simulate new observations for a given set of predictor values. This includes both uncertainty in the mean and the beta observation distribution. Below, I picked 2 arbitrary values for `x` to simulate with.

```{r}
newdata <- data.frame(x = 3)
p <- posterior_predict(fit, draws = 200, newdata = newdata)
hist(p)

newdata <- data.frame(x = -0.5)
p <- posterior_predict(fit, draws = 200, newdata = newdata)
hist(p)
```

We can also take draws from the posteriors:

```{r}
intercept <- as.matrix(fit)[,"(Intercept)"]
beta_x <- as.matrix(fit)[,"x"]

x_seq <- seq(min(x), max(x), length.out = 100)
plot(x_seq, plogis(coef(fit)[[1]] + coef(fit)[[2]] * x_seq), type = "l",
  ylim = c(0.5, 0.95), col = "red", lwd = 3, ylab = "Proportion")
for (i in seq_len(100)) {
  lines(x_seq, plogis(intercept[i] + beta_x[i] * x_seq), col = "#00000040")
}
```

# betareg

Now let's do something similar with the betareg package.

First I extracted the function `arm::sim()` from <https://github.com/cran/arm/blob/master/R/sim.R>

Then I modified it as needed to match the output from the betareg package:

```{r}
sim_betareg <- function(object, n.sims = 100) {
  object.class <- class(object)[[1]]
  summ <- summary (object, correlation=TRUE, dispersion = object$dispersion)
  coef <- summ$coefficients$mean[,1:2,drop=FALSE]
  dimnames(coef)[[2]] <- c("coef.est","coef.sd")
  beta.hat <- coef[,1,drop=FALSE]
  sd.beta <- coef[,2,drop=FALSE]
  k <- length(object$coefficients$mean)
  corr.beta <- cov2cor(vcov(object))[1:k, 1:k]
  V.beta <- corr.beta * array(sd.beta,c(k,k)) * t(array(sd.beta,c(k,k)))
  beta <- array (NA, c(n.sims,k))
  dimnames(beta) <- list (NULL, dimnames(beta.hat)[[1]])
  for (s in 1:n.sims){
    beta[s,] <- MASS::mvrnorm (1, beta.hat, V.beta)
  }
  beta2 <- array (0, c(n.sims,length(coefficients(object))))
  dimnames(beta2) <- list (NULL, names(coefficients(object)))
  beta2[,dimnames(beta2)[[2]]%in%dimnames(beta)[[2]]] <- beta
  beta2 <- beta2[, 1:k]
  beta2
}

```

Now we can fit the model and take draws from the coefficients.

```{r}
fit2 <- betareg(y ~ x, data = fake_dat, link = "logit")
summary(fit2)

s <- sim_betareg(fit2, n.sims = 100)

x_seq <- seq(min(x), max(x), length.out = 100)
plot(x_seq, plogis(coef(fit2)[[1]] + coef(fit2)[[2]] * x_seq), type = "l",
  ylim = c(0.5, 0.95), col = "red", lwd = 3, ylab = "Proportion")
for (i in seq_len(nrow(s))) {
  lines(x_seq, plogis(s[i, 1] + s[i, 2] * x_seq), col = "#00000040")
}
```

I think it's a little bit trickier if we also want to simulate new observations with the beta distribution because there is some correlation between the mean and the dispersion parameter.
